The brightness and spatial distributions of terrestrial radio sources

### 2013-05-21 articles/Article-RFI-distribution/article-rfi-distribution.tex
SUBMITTED

\begin{abstract}
Faint undetected sources of radio-frequency interference (RFI) might become visible after \editmark{long integrations} of radio observations if they \editmark{are consistently} present over time. Thereby, they might obstruct the detection of the weak astronomical signals of interest. This issue is especially important for Epoch of Reionisation (EoR) projects that try to detect the faint redshifted HI signals from the time of the earliest structures in the Universe. We explore the RFI situation \editmark{at 30--163~MHz} by studying brightness histograms of visibility data observed with LOFAR, similar to radio-source-count analyses that are used in cosmology. The histograms show an RFI distribution that follows a power-law distribution with an estimated exponent around -1.5. With several assumptions, this can be explained with a uniform distribution of terrestrial radio sources whose radiation follows existing propagation models. Extrapolation of the power law \editmark{implies} that the current LOFAR EoR observations should be severely RFI limited if the \editmark{strength of RFI sources remains strong after time integration}. This is in contrast with actual observations, which almost reach the thermal noise and are thought not to be limited by RFI. Therefore, we conclude that it is \editmark{unlikely that there are undetected RFI sources that will become visible after long integration. Consequently, there is no indication that RFI will prevent an EoR detection with LOFAR.}
\end{abstract}

>>>>

\begin{abstract}
Faint undetected sources of radio-frequency interference (RFI) might become visible \editmark{in long radio observations when} they \editmark{are consistently} present over time. Thereby, they might obstruct the detection of the weak astronomical signals of interest. This issue is especially important for Epoch of Reionisation (EoR) projects that try to detect the faint redshifted HI signals from the time of the earliest structures in the Universe. We explore the RFI situation \editmark{at 30--163~MHz} by studying brightness histograms of visibility data observed with LOFAR, similar to radio-source-count analyses that are used in cosmology. \editmark{An empirical RFI distribution model is derived that allows the simulation of RFI in radio observations.} The brightness histograms show an RFI distribution that follows a power-law distribution with an estimated exponent around -1.5. With several assumptions, this can be explained with a uniform distribution of terrestrial radio sources whose radiation follows existing propagation models. Extrapolation of the power law \editmark{implies} that the current LOFAR EoR observations should be severely RFI limited if the \editmark{strength of RFI sources remains strong after time integration}. This is in contrast with actual observations, which almost reach the thermal noise and are thought not to be limited by RFI. Therefore, we conclude that it is \editmark{unlikely that there are undetected RFI sources that will become visible in long observations. Consequently, there is no indication that RFI will prevent an EoR detection with LOFAR.}
\end{abstract}
